## Question 4: Are there differences in activity patterns between weekdays and weekends?  

### Q4-1. Create a new factor variable in the dataset with two levels: weekday and weekend - indicating whether a given date is a weekday or weekend day.  

```{r, dataset_with_weekdays, echo = TRUE}
dat <- read.csv(file = "./activity.csv") # with stringsAsFactors = FALSE
ave.steps.by.intv <- aggregate(steps ~ interval, data = dat,
                               function(x) mean(x, na.rm = TRUE))

# rename data frame so it makes easier to understand operations onwards.
target.df <- dat
new.val.df <- ave.steps.by.intv

dat.d <- dat[is.na(dat$steps), ]$date
lv <- levels(droplevels(dat.d))

# fill with an averege steps for NA's interval for each NA's date.
for (ymd in lv) { # NA's date: "2012-10-01", "2012-10-08"..
    tmp.df <- target.df[target.df$date == ymd, ]
    for (i in new.val.df$interval) { # 0, 5, 10, ... 2355
        new.v <- new.val.df[ new.val.df$interval == i, ]$steps
        target.df[target.df$date == ymd & target.df$interval == i, ]$steps <- new.v
    }
}

# up to here the code is the same as the one in Q3.

# create POSIXlt data
target.df$date.p <- as.POSIXlt(as.Date(as.character(target.df$date)))
# 0 is Sunday, 6 is Saturday
target.df$is.weekend <- target.df$date.p$wday == 0 | target.df$date.p$wday == 6

# split data by is.weekend (TRUE/FALSE)
sp.wd <- split(target.df, target.df$is.weekend)

ave.steps.by.intv.wd <- aggregate(steps ~ interval, data = sp.wd[[1]],
                                  function(x) mean(x, na.rm = TRUE))
ave.steps.by.intv.we <- aggregate(steps ~ interval, data = sp.wd[[2]],
                                  function(x) mean(x, na.rm = TRUE))

# now create data.frame for each of weekday, and weekend, then union them by rbind.
wd <- data.frame(week = rep("weekday", nrow(ave.steps.by.intv.wd)), ave.steps.by.intv.wd)
we <- data.frame(week = rep("weekend", nrow(ave.steps.by.intv.we)), ave.steps.by.intv.we)

ave.steps.by.intv.g <- rbind(wd, we)


str(ave.steps.by.intv.g)
```

The column **week** in ave.steps.by.intv.g data.frame is a new factor variable with two levels of weekday and weekend.  


### Q4-2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).   

Here goes a panel plot!  

```{r, panel_plot, echo = TRUE, fig.height = 10, fig.width = 12}
library(ggplot2)
g <- ggplot(data = ave.steps.by.intv.g,
            aes(x = interval, y = steps, group = week, color = week))
g <- g + geom_line()
g <- g + facet_grid(week ~ .)
g <- g + xlab("Interval")
g <- g + ylab("Number of steps")
g <- g + ggtitle("Average number of steps of 5-minutes interval\nfor weekday and weekend")
print(g)
```
